Tackling discrepancies in freshwater invertebrate trait databases and taxonomic resolution: Harmonising across continents and trait aggregation

### 2021-01-27 Paper/Draft_harmonization_paper.tex
27.01: Minor fixes before send to Brooke

1. Use of invertebrate traits rather than species composition is expected to facilitate large-scale comparisons of community structure and responses to disturbance in freshwater ecology because the same traits can potentially occur everywhere. % OLD: because of their consistent response to environmental gradients
In recent years, comprehensive invertebrate trait databases have been established at different scales (e.g. regions, continents). The wide availability of invertebrate trait data supports trait-based studies, especially at large scales. However, a number of data-related issues complicate the use of invertebrate traits for ecological studies. For example, for freshwater invertebrate traits standardised definitions are lacking, impeding comparisons across regions. Moreover, it is uncertain how harmonising varying trait definitions between databases might influence trait-environment relationships. Another issue is that if aggregation of traits is necessary, i.e. when observational taxonomic data is less precise than available trait data, no guidance exists which method under which circumstances can 
% Alternative from Co-Authors: Another issue is that we do not know how taxonomic resolution affects the characterization of traits and hence the comparability of different datasets.

2. We describe discrepancies in the definitions of traits used to create freshwater invertebrate trait databases in Europe, North America, New Zealand, and Australia. Based on our comparisons of these trait databases, we established 4 novel trait datasets by harmonising trait definitions of commonly used traits. Two of these datasets were used to compare aggregated traits obtained by different aggregation methods with traits assigned by experts at family-level. We compared trait aggregation methods that use either the mean or the median and different weightings. Furthermore, we explored the effect of harmonisation and trait aggregation by re-analysing data from a case study.

3. Between the trait databases, grouping features differed in terms of their classification into traits and their focus leading to discrepancies how traits are defined. Also, the coding to describe traits varied between databases. Aggregated and assigned traits based on expert knowledge displayed a good match, especially approaches using the median to aggregate from species to family-level. Harmonisation and trait aggregation only slightly changed the results in the re-analysis compared to the original study, although averaging effects did occur in some cases because of harmonisation.

@@

To our knowledge, only \citet{brown_functional_2018} harmonised grouping features from more than two geographically distant invertebrate trait databases, for a limited set of grouping features (8) and taxa (112), in a study on the influence of decreasing glacier cover on functional diversity and community assembly of invertebrates. We suspect that the heterogeneity of information in freshwater invertebrate trait databases, besides the diversity of taxa across regions, is likely a major reason for the lack of studies across continents. To harmonize grouping features from different regions, first commonly accepted and unambiguous trait definitions are required (\cite{schneider_towards_2019}). In the best case, grouping features would be classified into the same traits across databases or they could easily be harmonised using standardised terminology. However, a lack of standardised terminology of trait definitions and poor metadata quality in many trait databases are common issues throughout the field of trait-based ecology (\cite{baird_toward_2011, kissling_towards_2018}). Secondly, consistent coding of traits facilitates the compatibility of trait data from different databases.

@@

However, invertebrate trait databases are heterogeneous regarding the coding they use for their traits (\cite{culp_incorporating_2011}). For example, in the study from \citet{brown_functional_2018} the authors needed to reclassify traits using expert knowledge given that the individual databases employed different coding approaches (i.e. European and New Zealand databases fuzzy coding, North American database binary coding). Databases with standardised and unambiguous traits and a consistent coding of traits would minimize the data processing effort. Discrepancies in the taxonomic resolutions (e.g. species, genus or family) when linking observational taxonomic data to trait databases represents another challenge. When observations are at a more precise taxonomic level than data available in the trait databases (e.g. observations at species-level, trait data at genus-level) trait information of the less precise taxonomic level is often assigned (e.g. \cite{szocs_effects_2014, vos_taxonomic_2017}). Conversely, if trait information is only available at more precise taxonomic levels than the observed taxa, traits are aggregated to a less precise taxonomic level (e.g. \cite{aspin_extreme_2019, piliere_a._f._h._importance_2016, poff_functional_2006, szocs_effects_2014}). To data, studies have used different methods of trait aggregation, e.g. the mean (\cite{magliozzi_functional_2019}), median (\cite{szocs_effects_2014}) or the mode (\cite{piliere_a._f._h._importance_2016}). Studies on how and to which extent different trait aggregation methods influence trait-based analysis are missing. However, related knowledge would inform future studies regarding the choice of the aggregation method.

These challenges when working with or synthesizing trait data motivated us to (1) investi<gate discrepancies in trait definition between trait databases from Europe, North America, Australia, and New Zealand. As a basis for further evaluating different trait aggregation methods and how harmonising definition discrepancies between grouping features can influence trait-environment relationships, we developed 4 novel invertebrate trait datasets from these existing trait databases by harmonising 7 grouping features. (2) We then evaluated different trait aggregation methods by comparing aggregated traits with traits assigned by experts at family-level for two of the established trait datasets. (3) Finally, we re-analysed data from a study of salinisation effects on biological traits (\cite{szocs_effects_2014}) using data from the newly established harmonised European trait dataset and aggregated traits. By comparing the re-analysis's findings with those of \citet{szocs_effects_2014}, we investigated how harmonising and aggregating trait data can modify the outcome of trait-environment relationships.

@@

We repeated the analysis of \citet{szocs_effects_2014} who studied the effect of anthropogenic salinization on invertebrates in the River Werra in Germany. For this re-analysis we used the established harmonised grouping feature dataset for Europe and additionally aggregated traits using the aforementioned aggregation methods.

The River Werra has been subject to effluents from the potash industry since the mid- 20th century and allows to study responses of invertebrates and their trait compositions to salinization (\cite{bathe_biological_2011}). Sites downstream, upstream, and close to the salt discharge (transition) were compared regarding their trait composition. Further details can be found in \citet{szocs_effects_2014}.

We substituted 6 of the grouping features from the original data with harmonised grouping features from the European trait dataset. We compared the results of the redundancy analysis (RDA) from the original study to the case when using harmonised grouping features. Specifically, the trait composition expressed as community weighted mean (CWM) traits was ordinated along an electric conductivity gradient. We compared the species scores obtained from the RDA, i.e. the coordinates of the tips of the vectors representing the CWM traits in the bi- or triplots. Following the original study, we identified traits associated with high or low salinity based on their distance to the ordination axis median using the Mahalanobis distance. CWM traits with a Mahalanobis distance greater than the $97.5 \%$ -quantile of the Chi-square distribution (5.02) were regarded as traits responding to either low or high salinity. For our analysis, we used the same 21 grouping features as in \citet{szocs_effects_2014}. The 6 harmonised grouping features used were \textit{Size, Feeding mode, Locomotion, Oviposition, Respiration}, and \textit{Voltinism}. Additionally, for testing the effect of aggregated traits we assigned to each taxon in \citet{szocs_effects_2014} the aggregated trait value for its corresponding family and repeated the RDA.

@@

The definition of grouping features varied across databases mainly concerning their differentiation into traits but also in their focus. We provide a summary of discrepancies in trait definitions in the appendix (Table \ref{stab:trait_definitions}). Both, differences in differentiation and scope can lead to discrepancies in trait definitions. For example, for the grouping feature feeding mode discrepancies arise because traits are assigned in different ways. The Tachet database defines predators as carvers, engulfers and swallowers. By contrast, in the CONUS database predators are defined as engulfers and carnivorous piercers. In turn, in the Tachet database, piercers are defined as a separate trait encompassing herbivorous and carnivorous piercers. Furthermore, the focus in the freshwaterecology.info database for feeding mode is primarily on the food source of a species (except for filterers), while the other databases focus on the strategies of food acquisition. Therefore, the freshwaterecology.info database defines e.g. predator as "eating from prey", while the other databases use the mouthpart morphology as basis of their definition. The Tachet database captures the food source in an additional grouping feature. Varying levels of differentiation are also present in all other investigated grouping features between the trait databases (Table \ref{tab:trait_databases_coding_differentiation} and Table \ref{stab:trait_definitions}). Locomotion definitions differ also in focus between databases. Freshwaterecology.info and New Zealand databases describe locomotion as the way of movement of an organism, Tachet as substrate relation and locomotion, the Vieira as how organisms deal with flow, Australia as attachment, and the CONUS database includes not only the way of movement, but also the location of movement. Similarly, regarding reproduction traits, databases differ in their focus. Reproduction is captured in one grouping feature and defined as location of oviposit clutches and mode of reproduction in the freshwaterecology.info and Tachet databases. The Vieira database provides information on the oviposition location but not on reproductive behavior. The Australian database report traits for reproductive behavior but also on oviposition site. The New Zealand database distinguishes three grouping features related to reproduction: reproductive technique, oviposition site (e.g. water surface, terrestrial), and egg/egg mass (e.g. free, cemented).

@@

%BEN: Suggest start with some overall main message. The main message I get from Fig 6 is: „The original analysis in Szocs et al. (2014) and all agrigration methods produced very similar distributions of RDA species scores (Figure 6).“ I suggest start with this sentence (or something like this), then  draw out any details but only if necessary (i.e. do not feel compelled to say more if there is not really much more to say).

Using harmonised grouping features resulted in less traits that distinguish upstream and downstream sites in comparison to the original analysis (Figures \ref{fig:violin_plot_species_sc} and \ref{fig:boxplots_scores_on_constrained_axis}). According to the RDA of the trait composition, downstream sites were characterised by taxa with the traits multivoltinism, ovoviviparity and long life cycle duration. Upstream sites were characterised by univoltine taxa with a short life cycle duration that lay their eggs in an aquatic environment (aquatic eggs). The trait aquatic eggs of the harmonised grouping feature oviposition was derived by amalgamating the trait oviposition in clutches and other related traits (Table \ref{tab:traits_harmonisation}). The traits shredder and gill respiration did not characterise sites with high salinisation.

Using traits aggregated at family-level from the harmonised dataset also showed results similar to the original analysis, but with fewer traits distinguishing upstream and downstream sites (Figures \ref{fig:violin_plot_species_sc} and \ref{fig:boxplots_scores_on_constrained_axis}). The \textit{direct\_agg\textsubscript{mean}}, \textit{direct\_agg\textsubscript{median}}, and \textit{weighted\_agg} 
characterised the downstream sites with the same traits as the original analysis except that downstream sites were not characterised by the trait shredder. Upstream sites were characterised by the traits univoltinism, aquatic eggs and short life cycle duration. The same results were obtained when re-analysing the data with traits aggregated by the \textit{stepwise\_agg\textsubscript{mean}} and \textit{stepwise\_agg\textsubscript{median}}, with the exception that none of the life cycle traits characterised upstream or downstream sites. Thus, in our comparison aggregating by \textit{direct\_agg\textsubscript{mean}}, \textit{direct\_agg\textsubscript{median}}, and \textit{weighted\_agg} yielded the least change in interpretation of the RDA results compared to the original findings.

@@

Although comprehensive freshwater invertebrate trait databases have been developed for several regions in the past, data synthesis is difficult due to discrepancies in trait definitions. By providing an overview of these definition discrepancies we provide a starting point for the development of standardised trait terminology through which invertebrate trait databases can be linked. A consent on standard terminology and the subsequent development of ontologies are the next steps to facilitate trait-based analysis at large geographical scales. As our analysis showed, some grouping features might need to be re-classified to fit into such a standardised terminology. We could show that trait affinities resulting from fuzzy coding and binary coding can be used together, but a uniform coding of traits is another problem that should be addressed during trait standardisation.

As ecological knowledge increases in the future, and with the aid of new  methods such as computer vision for species and trait identification (\cite{hoye_deep_2020}), more observational and trait data at species-level may become available in the future, reducing the need for trait aggregation. However, if traits need to be aggregated we showed that trait aggregation is often in agreement with expert assignments, especially aggregation approaches using the median. Ultimately, the question which aggregation method is the objectively best measure cannot be answered. Nevertheless, we make the following recommendations for situations when trait aggregation is necessary 1) median aggregation seems to be favourable, 2) highly variable traits and taxa where a few genera have a high number of species should be treated carefully, 3) description of traits (binary, fuzzy) should be considered, 4) weighting trait information based on the taxa present in the database used does not seem to have much influence and 5) a measure of trait dispersion should be reported to indicate the uncertainty of the aggregated estimate. 
% OLD: No matter which aggregation approach is chosen, for future studies in which researchers must aggregate traits,....

>>>>

1. Use of invertebrate traits rather than species composition is expected to facilitate large-scale comparisons of community structure and responses to disturbance in freshwater ecology because the same traits can potentially occur everywhere. % OLD: because of their consistent response to environmental gradients
In recent years, comprehensive invertebrate trait databases have been established at different scales (e.g. regions, continents). The wide availability of invertebrate trait data supports trait-based studies, especially at large scales. However, a number of data-related issues complicate the use of invertebrate traits for ecological studies. For example, for freshwater invertebrate traits standardised definitions are lacking, impeding comparisons across regions. Moreover, it is uncertain how harmonising varying trait definitions between databases might influence trait-environment relationships. Another issue is that taxonomic resolution between observational taxonomic data and trait data can differ, making aggregation of traits necessary. So far, comparisons of different aggregation approaches to expert assessments are lacking. 
% Alternative from Co-Authors: Another issue is that we do not know how taxonomic resolution affects the characterization of traits and hence the comparability of different datasets.

2. We describe discrepancies in the definitions of traits used to create freshwater invertebrate trait databases in Europe, North America, New Zealand, and Australia. Based on our comparisons of these trait databases, we established 4 novel trait datasets by harmonising trait definitions of commonly used traits. Two of these datasets were used to compare aggregated traits obtained by different aggregation methods with traits assigned by experts at family-level. We compared trait aggregation methods that use either the mean or the median and different weightings. Thereby, we simulated how taxonomic resolution and trait variability can change the result of different aggregation approaches.
Furthermore, we explored the effect of harmonisation and trait aggregation by re-analysing data from a case study.

% Intial: Between the trait databases, grouping features differed in terms of their classification into traits and their focus leading to discrepancies how traits are defined
3. We found that between trait databases, trait definitions are often different because varying numbers of traits are used to describe groups of related traits (e.g. respiration traits). Also the focus to describe groups of related traits varies between databases, e.g. for feeding mode some focus on the food source, while others on mouthpart morphology. Also, the coding to describe traits (binary, fuzzy) varied between databases. Aggregated and assigned traits based on expert knowledge displayed a good match, especially approaches using the median to aggregate from species to family-level. The comparison with the case study showed that harmonisation and trait aggregation only slightly changed results.

@@

To our knowledge, only \citet{brown_functional_2018} harmonised grouping features from more than two geographically distant invertebrate trait databases, for a limited set of grouping features (8) and taxa (112), in a study on the influence of decreasing glacier cover on functional diversity and community assembly of invertebrates. We suspect that the heterogeneity of information in freshwater invertebrate trait databases, besides the diversity of taxa across regions, is likely a major reason for the lack of studies across continents. To harmonize grouping features from different regions, first commonly accepted and unambiguous trait definitions are required (\cite{schneider_towards_2019}). In the best case, grouping features would be classified into the same traits across databases or they could easily be harmonised using standardised terminology. However, a lack of standardised terminology of trait definitions and poor metadata quality in many trait databases are common issues throughout the field of trait-based ecology (\cite{baird_toward_2011, kissling_towards_2018}). Secondly, consistent coding of traits facilitates the comparability of trait data from different databases.

@@

However, invertebrate trait databases are heterogeneous with respect to the coding that is used for traits (\cite{culp_incorporating_2011}). For example, in the study from \citet{brown_functional_2018} the authors needed to reclassify traits using expert knowledge given that the individual databases employed different coding approaches (i.e. European and New Zealand databases fuzzy coding, North American database binary coding). Databases with standardised and unambiguous traits and a consistent coding of traits would minimize the data processing effort. Discrepancies in the taxonomic resolutions (e.g. species, genus or family) when linking observational taxonomic data to trait databases represents another challenge. When observations are at a more precise taxonomic level than data available in the trait databases (e.g. observations at species-level, trait data at genus-level) trait information of the less precise taxonomic level is often assigned (e.g. \cite{szocs_effects_2014, vos_taxonomic_2017}). Conversely, if trait information is only available at more precise taxonomic levels than the observed taxa, traits are aggregated to a less precise taxonomic level (e.g. \cite{aspin_extreme_2019, piliere_a._f._h._importance_2016, poff_functional_2006, szocs_effects_2014}). To data, studies have used different methods of trait aggregation, e.g. the mean (\cite{magliozzi_functional_2019}), median (\cite{szocs_effects_2014}) or the mode (\cite{piliere_a._f._h._importance_2016}). Studies on how and to which extent different trait aggregation methods influence trait-based analysis are missing. However, related knowledge would inform future studies regarding the choice of the aggregation method.

These challenges when working with or synthesizing trait data motivated us to (1) investigate discrepancies in trait definition between trait databases from Europe, North America, Australia, and New Zealand. As a basis for further evaluating different trait aggregation methods and how harmonising definition discrepancies between grouping features can influence trait-environment relationships, we developed 4 novel invertebrate trait datasets from these existing trait databases by harmonising 7 grouping features. (2) We then evaluated different trait aggregation methods by comparing aggregated traits with traits assigned by experts at family-level for two of the established trait datasets. (3) Finally, we re-analysed data from a study of salinisation effects on biological traits (\cite{szocs_effects_2014}) using data from the newly established harmonised European trait dataset and aggregated traits. By comparing the re-analysis's findings with those of \citet{szocs_effects_2014}, we investigated how harmonising and aggregating trait data can modify the outcome of trait-environment relationships.

@@

We repeated the analysis of \citet{szocs_effects_2014} who studied the effect of anthropogenic salinisation on invertebrates in the River Werra in Germany. For this re-analysis we used the established harmonised grouping feature dataset for Europe and additionally aggregated traits using the aforementioned aggregation methods.

The River Werra has been subject to effluents from the potash industry since the mid- 20th century and allows to study responses of invertebrates and their trait compositions to salinisation (\cite{bathe_biological_2011}). Sites downstream, upstream, and close to the salt discharge (transition) were compared regarding their trait composition. Further details can be found in \citet{szocs_effects_2014}.

We substituted 6 of the grouping features from the original data with harmonised grouping features from the European trait dataset. We compared the results of the redundancy analysis (RDA) from the original study to the case when using harmonised grouping features. Specifically, the trait composition expressed as community weighted mean (CWM) traits was ordinated along an electric conductivity gradient. We compared the species scores obtained from the RDA, i.e. the coordinates of the tips of the vectors representing the CWM traits in the bi- or triplots. Following the original study, we identified traits associated with higher or lower salinity based on their distance to the ordination axis median using the Mahalanobis distance. CWM traits with a Mahalanobis distance greater than the $97.5 \%$ -quantile of the Chi-square distribution (5.02) were regarded as traits responding to either lower or higher salinity. For our analysis, we used the same 21 grouping features as in \citet{szocs_effects_2014}. The 6 harmonised grouping features used were \textit{Size, Feeding mode, Locomotion, Oviposition, Respiration}, and \textit{Voltinism}. Additionally, for testing the effect of aggregated traits we assigned to each taxon in \citet{szocs_effects_2014} the aggregated trait value for its corresponding family and repeated the RDA.

@@

The definition of grouping features varied across databases mainly concerning their differentiation into traits but also in their focus. We provide a summary of discrepancies in trait definitions in the appendix (Table \ref{stab:trait_definitions}). Both, differences in differentiation and focus can lead to discrepancies in trait definitions. For example, for the grouping feature feeding mode discrepancies arise because traits are assigned in different ways. The Tachet database defines predators as carvers, engulfers and swallowers. By contrast, in the CONUS database predators are defined as engulfers and carnivorous piercers. In turn, in the Tachet database, piercers are defined as a separate trait encompassing herbivorous and carnivorous piercers. Furthermore, the focus in the freshwaterecology.info database for feeding mode is primarily on the food source of a species (except for filterers), while the other databases focus on the strategies of food acquisition. Therefore, the freshwaterecology.info database defines e.g. predator as "eating from prey", while the other databases use the mouthpart morphology as basis of their definition. The Tachet database captures the food source in an additional grouping feature. Varying levels of differentiation are also present in all other investigated grouping features between the trait databases (Table \ref{tab:trait_databases_coding_differentiation} and Table \ref{stab:trait_definitions}). Locomotion definitions differ also in focus between databases. Freshwaterecology.info and New Zealand databases describe locomotion as the way of movement of an organism, Tachet as substrate relation and locomotion, the Vieira as how organisms deal with flow, Australia as attachment, and the CONUS database includes not only the way of movement, but also the location of movement. Similarly, regarding reproduction traits, databases differ in their focus. Reproduction is captured in one grouping feature and defined as location of oviposit clutches and mode of reproduction in the freshwaterecology.info and Tachet databases. The Vieira database provides information on the oviposition location but not on reproductive behavior. The Australian database report traits for reproductive behavior but also on oviposition site. The New Zealand database distinguishes three grouping features related to reproduction: reproductive technique, oviposition site (e.g. water surface, terrestrial), and egg/egg mass (e.g. free, cemented).

@@

Overall, harmonised trait data and all aggregation methods produced similar distributions of RDA species scores (Figures \ref{fig:violin_plot_species_sc}) compared to the original analysis in \citet{szocs_effects_2014}. Using harmonised grouping features resulted in less traits that distinguish upstream and downstream sites in comparison to the original analysis (Figures \ref{fig:violin_plot_species_sc} and \ref{fig:boxplots_scores_on_constrained_axis}). According to the RDA of the trait composition, downstream sites were characterised by taxa with the traits multivoltinism, ovoviviparity and long life cycle duration. Upstream sites were characterised by univoltine taxa with a short life cycle duration that lay their eggs in an aquatic environment (aquatic eggs). The trait aquatic eggs of the harmonised grouping feature oviposition was derived by amalgamating the trait oviposition in clutches and other related traits (Table \ref{tab:traits_harmonisation}). The traits shredder and gill respiration did not characterise sites with higher salinisation.

Using traits aggregated at family-level from the harmonised dataset also showed results similar to the original analysis, but with less traits distinguishing upstream and downstream sites (Figures \ref{fig:violin_plot_species_sc} and \ref{fig:boxplots_scores_on_constrained_axis}). The \textit{direct\_agg\textsubscript{mean}}, \textit{direct\_agg\textsubscript{median}}, and \textit{weighted\_agg} 
characterised the downstream sites with the same traits as the original analysis except that downstream sites were not characterised by the trait shredder. Upstream sites were characterised by the traits univoltinism, aquatic eggs and short life cycle duration. The same results were obtained when re-analysing the data with traits aggregated by the \textit{stepwise\_agg\textsubscript{mean}} and \textit{stepwise\_agg\textsubscript{median}}, with the exception that none of the life cycle traits characterised upstream or downstream sites. Thus, in our comparison aggregating by \textit{direct\_agg\textsubscript{mean}}, \textit{direct\_agg\textsubscript{median}}, and \textit{weighted\_agg} yielded the least change in interpretation of the RDA results compared to the original findings.

@@

Although comprehensive freshwater invertebrate trait databases have been developed for several regions in the past, data synthesis is difficult because of discrepancies in trait definitions. By providing an overview of these definition discrepancies we have set a starting point for the development of standardised trait terminology through which invertebrate trait databases can be linked. A consent on standard terminology and the subsequent development of ontologies are the next steps to facilitate trait-based analysis at large geographical scales. As our analysis showed, some grouping features might need to be re-classified to fit into such a standardised terminology. We could show that trait affinities resulting from fuzzy coding and binary coding can be used together, but a uniform coding of traits is another problem that should be addressed during trait standardisation.

As ecological knowledge increases in the future, and with the aid of new  methods such as computer vision for species and trait identification (\cite{hoye_deep_2020}), more observational and trait data at species-level may become available in the future, reducing the need for trait aggregation. However, if traits need to be aggregated we showed that trait aggregation is often in agreement with expert assignments, especially aggregation approaches using the median. Ultimately, the question which aggregation method is the objectively best measure cannot be answered. Nevertheless, we make the following recommendations for situations when trait aggregation is necessary. 1) The aggregation method will tend to have little effect except for highly variable traits and taxa where a few genera have a high number of species. 2) The coding used to describe traits (binary, fuzzy) should be considered, because median and mean aggregations can yield very different results for binary coded traits. 3) Weighting trait information based on the taxa present in the database used does not seem to have much influence except for families with few genera that have a high number of species. 4) A measure of trait dispersion should be reported to indicate the uncertainty of the aggregated estimate. 
% OLD: No matter which aggregation approach is chosen, for future studies in which researchers must aggregate traits,....

### 2021-01-27 Paper/Results/SI.tex
22.12: Slight changes in SI to avoid merging conflict
27.01: Minor fixes before send to Brooke

    \begin{longtable}{m{1.8cm}|m{3cm}|m{3cm}|m{3cm}|m{3cm}|m{3cm}|m{3cm}}
        \caption{Comparison of trait definitions between invertebrate trait databases. Only traits that are differently described across databases are listed. The definition is quoted if it enables differences to be identified, otherwise the differences are described. The hyphen indicates a missing trait. Reproduction was captured in multiple grouping features per database. Hence, differences for reproduction have been described in the paper. Body form traits are not different between databases, except that the North America (Vieira et al.) database contains the trait Bluff (blocky) which does not appear in the other databases.}

@@

        Trait & \specialcell{Freshwater- \\ ecology.info} & Tachet & \specialcell{North America \\ (Twardochleb et al.)} & 
        North America (Vieira et al.) & Australia & New Zealand \\

@@

\subsection*{Re-analysis of Szöcs et al. using harmonized and aggregated grouping features}

\begin{table}[ht]
    \centering
    \caption{Mean, median and standard deviation of trait affinities that were responsive to the salinity gradient in the original study but not when re-analysis with the harmonized dataset trait dataset.} 
    \label{tab:SI_resp_traits_summary_stats}
    \begin{tabular}{llccc}
    \toprule[.1em]
    Dataset & Trait & Mean & Median & SD \\ 
    \toprule[.1em]
    Stepw\_median & Shredder & 0.20 & 0.14 & 0.25 \\ 
      Stepw\_mean & Shredder & 0.18 & 0.12 & 0.22 \\ 
      Direct\_median & Shredder & 0.21 & 0.14 & 0.25 \\ 
      Direct\_mean & Shredder & 0.19 & 0.14 & 0.22 \\ 
      Weighted & Shredder & 0.19 & 0.14 & 0.22 \\ 
      Harmonized; not\_aggregated & Shredder & 0.18 & 0.12 & 0.24 \\ 
      Original & Shredder & 0.25 & 0.14 & 0.32 \\ 
      Stepw\_median & Gills & 0.30 & 0.27 & 0.32 \\ 
      Stepw\_mean & Gills & 0.29 & 0.22 & 0.32 \\ 
      Direct\_median & Gills & 0.30 & 0.30 & 0.32 \\ 
      Direct\_mean & Gills & 0.30 & 0.30 & 0.32 \\ 
      Weighted & Gills & 0.30 & 0.30 & 0.32 \\ 
      Harmonized; not\_aggregated & Gills & 0.30 & 0.25 & 0.32 \\ 
      Original & Gills & 0.28 & 0.00 & 0.33 \\ 
      Stepw\_median & Short life cycle & 0.64 & 0.75 & 0.39 \\ 
      Stepw\_mean & Short life cycle & 0.64 & 0.79 & 0.39 \\ 
      Direct\_median & Short life cycle & 0.67 & 0.75 & 0.37 \\ 
      Direct\_mean & Short life cycle & 0.67 & 0.79 & 0.38 \\ 
      Weighted & Short life cycle & 0.67 & 0.79 & 0.38 \\ 
      Harmonized; not\_aggregated & Short life cycle & 0.64 & 0.75 & 0.40 \\ 
      Original & Short life cycle & 0.64 & 0.75 & 0.40 \\ 
      Stepw\_median & Long life cylce & 0.36 & 0.25 & 0.39 \\ 
      Stepw\_mean & Long life cylce & 0.36 & 0.21 & 0.39 \\ 
      Direct\_median & Long life cylce & 0.33 & 0.25 & 0.37 \\ 
      Direct\_mean & Long life cylce & 0.33 & 0.21 & 0.38 \\ 
      Weighted & Long life cylce & 0.33 & 0.21 & 0.38 \\ 
      Harmonized; not\_aggregated & Long life cylce & 0.36 & 0.25 & 0.40 \\ 
      Original & Long life cylce & 0.36 & 0.25 & 0.40 \\ 
    \bottomrule
    \end{tabular}
\end{table}

>>>>

    \begin{longtable}{m{1.8cm}|m{3cm}|m{3cm}|m{3cm}|m{3cm}|m{3.2cm}|m{3cm}}
        \caption{Comparison of trait definitions between invertebrate trait databases. Only traits that are differently described across databases are listed. The definition is quoted if it enables differences to be identified, otherwise the differences are described. The hyphen indicates a missing trait. Reproduction was captured in multiple grouping features per database. Hence, differences for reproduction have been described in the paper. Body form traits are not different between databases, except that the Vieira database contains the trait Bluff (blocky) which does not appear in the other databases.}

@@

        Trait & \specialcell{Freshwater- \\ ecology.info} & Tachet & CONUS & Vieira & Australia & New Zealand \\

@@

\subsection*{Re-analysis of Szöcs et al. using harmonised and aggregated grouping features}

\begin{table}[ht]
    \centering
    \caption{Mean, median and standard deviation of the affinities of traits that were responsive to the salinity gradient in the original study but not in the re-analysis using the harmonised European trait dataset.} 
    \label{tab:SI_resp_traits_summary_stats}
    \begin{tabular}{llccc|c}
    \toprule[.1em]
    Type & Trait & Mean & Median & SD & Responsive? \\ 
    \toprule[.1em]
    Stepw\_median & Shredder & 0.20 & 0.14 & 0.25 & No \\ 
      Stepw\_mean & Shredder & 0.18 & 0.12 & 0.22 & No\\ 
      Direct\_median & Shredder & 0.21 & 0.14 & 0.25 & No\\ 
      Direct\_mean & Shredder & 0.19 & 0.14 & 0.22 & No\\ 
      Weighted & Shredder & 0.19 & 0.14 & 0.22 & No \\ 
      Harmonised; not\_aggregated & Shredder & 0.18 & 0.12 & 0.24 & No \\ 
      Original & Shredder & 0.25 & 0.14 & 0.32 & Yes\\ 
      \midrule
      Stepw\_median & Gills & 0.30 & 0.27 & 0.32 & Yes\\ 
      Stepw\_mean & Gills & 0.29 & 0.22 & 0.32 & Yes\\ 
      Direct\_median & Gills & 0.30 & 0.30 & 0.32 & Yes\\ 
      Direct\_mean & Gills & 0.30 & 0.30 & 0.32 & Yes\\ 
      Weighted & Gills & 0.30 & 0.30 & 0.32 & Yes\\ 
      Harmonised; not\_aggregated & Gills & 0.30 & 0.25 & 0.32 & No \\ 
      Original & Gills & 0.28 & 0.00 & 0.33 & Yes \\ 
      \midrule
      Stepw\_median & Short life cycle & 0.64 & 0.75 & 0.39 & No \\ 
      Stepw\_mean & Short life cycle & 0.64 & 0.79 & 0.39 & No \\ 
      Direct\_median & Short life cycle & 0.67 & 0.75 & 0.37 & Yes \\ 
      Direct\_mean & Short life cycle & 0.67 & 0.79 & 0.38 & Yes \\ 
      Weighted & Short life cycle & 0.67 & 0.79 & 0.38 & Yes\\ 
      Harmonised; not\_aggregated & Short life cycle & 0.64 & 0.75 & 0.40 & Yes \\ 
      Original & Short life cycle & 0.64 & 0.75 & 0.40 & Yes \\ 
      \midrule
      Stepw\_median & Long life cylce & 0.36 & 0.25 & 0.39 & No \\ 
      Stepw\_mean & Long life cylce & 0.36 & 0.21 & 0.39 & No \\
      Direct\_median & Long life cylce & 0.33 & 0.25 & 0.37 & Yes\\ 
      Direct\_mean & Long life cylce & 0.33 & 0.21 & 0.38 & Yes \\ 
      Weighted & Long life cylce & 0.33 & 0.21 & 0.38 & Yes \\ 
      Harmonised; not\_aggregated & Long life cylce & 0.36 & 0.25 & 0.40 & Yes \\ 
      Original & Long life cylce & 0.36 & 0.25 & 0.40 & Yes\\ 
    \bottomrule
    \end{tabular}
\end{table}
